### 2014-09-23 Studies/2012_Chicago/Report/Chicago_Fire.tex
San Francisco: Minor formatting edits
Chicago Fire: update to report to fit NIST TN style

\begin{document}

\bibliographystyle{unsrt}
\pagestyle{empty}

@@

\pagenumbering{roman}

@@

\begingroup
\hypersetup{urlcolor=black}
\href{http://dx.doi.org/10.6028/NIST.TN.1838}{http://dx.doi.org/10.6028/NIST.TN.1838} \\
\endgroup

@@

@@

\pagestyle{plain}
\pagenumbering{roman}

@@

\chapter*{\centering Abstract}
The Fire Dynamics Simulator (FDS) fire model, which is developed and maintained by the National Institute of Standards and Technology (NIST), was used to provide insight into the dynamics of a fire that occurred on November 2, 2012, within a 2~\sfrac{1}{2}~story, single family structure in Chicago, IL, that resulted in the death of the captain of a fire department. The inputs for the FDS simulations are documented in this report and are based on the fire scenario, including the building geometry, interior furnishings, and ventilation conditions. The fire started in the attic and spread down through an opening into the enclosed porch on the rear of the structure. This resulted in ventilation limited (fuel rich) fire conditions in the attic and rear porch areas. The spread of fire and hot gases into the second floor of the structure was limited by a closed steel-faced, wood-framed door. After exposure to the elevated temperatures and pressure from the fire in the porch area, the wood frame of the door decomposed, the steel faces of the door failed, and the door collapsed inward. The door failure resulted in the establishment of a flow path between the higher pressure and higher temperature conditions in the enclosed porch and the lower pressure and lower temperature conditions in the hallway and kitchen areas of the second floor. The temperature of the gases that flowed into the hallway exceeded 260~$^{\circ}$C (500~$^{\circ}$F) at a height of 1.86~m (6~ft). Unfortunately, a captain and firefighter were advancing a hose line in the hallway near the door at the time of the door failure. It appears that the captain was likely overcome by the rapid change in conditions before he could exit to the safety of the kitchen area. After a Mayday call, the captain was rescued from the structure, but later succumbed to his injuries.

\chapter{Introduction}
Part of the function of the Fire Research Division at NIST is to develop and apply technology, measurements, and standards to improve the understanding of the behavior, prevention, and control of fires to enhance fire fighting operations and equipment, fire suppression, fire investigations, and disaster response. NIST has previously used the Fire Dynamics Simulator (FDS) fire model to provide insight into the fire development and thermal conditions of fires that have resulted in line of duty deaths (LODDs)~\cite{Madrzykowski:1,Iowa,Texas,Bryner:Charleston,barowy:texas}. The objective of these studies has been to improve the safety and effectiveness of firefighters.

@@

\begin{table}
\centering
\captionof{table}{Relevant fire model input parameters}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter                                                  & Description                                   & Discussion                            \\
\midrule
Simulation Time                                            &  5~min                                        & --                                    \\ [.25cm]
Grid Cell Size                                             &  10~cm                                        & Section~\ref{mesh}                    \\ [.25cm]
Ambient Temperature$^*$                                    &  20~$^{\circ}$C (68~$^{\circ}$F)              & --                                    \\ [.1cm]
\multirow{4}{*}{Reaction: Wood~\cite{SFPE:Tewarson}}       &  Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$     & \multirow{4}{*}{Section~\ref{fire}}   \\
                                                           &  CO Yield: 0.004~$\mathrm{kg}/\mathrm{kg}$    &                                       \\
                                                           &  Soot Yield: 0.015~$\mathrm{kg}/\mathrm{kg}$  &                                       \\
                                                           &  Heat of Combustion: 16,400~kJ/kg             &                                       \\ [.25cm]
\multirow{2}{*}{Peak HRR}                                  &  Attic Fire: 9~MW                             & \multirow{2}{*}{Section~\ref{fire}}   \\ 
                                                           &  Porch Fire: 3.3~MW                           &                                       \\ [.25cm]                     
\multirow{3}{*}{Material: Wood~\cite{Incropera:1} }        &  $k$: 0.12~\si{W/(m.K)})                      & \multirow{3}{*}{Section~\ref{matl}}   \\
                                                           &  $\rho$: 510~kg/m$^3$                         &                                       \\
                                                           &  $c_{p}$: 1.38~\si{kJ/(kg.K)}                 &                                       \\ [.25cm]
\multirow{3}{*}{Material: Gypsum Board~\cite{WAKILI2007}}  &  $k$: 0.28~\si{W/(m.K)}                       &  \multirow{3}{*}{Section~\ref{matl}}  \\ 
                                                           &  $\rho$: 810~kg/m$^3$                         &                                       \\ 
                                                           &  $c_{p}$: 1.0~\si{kJ/(kg.K)}                  &                                       \\
\bottomrule[1.25pt]
\end{tabular}\par
\footnotesize
$^{*}$ Initial interior temperatures were assumed to be 20~$^{\circ}$C; exterior temperatures were 7~$^{\circ}$C~\cite{NIOSH:Bowyer}.
\normalsize
\end{table}

>>>>

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\rhead{}
\chead{This publication is available free of charge from:
\hypersetup{urlcolor=black}
\href{http://dx.doi.org/10.6028/NIST.TN.1838}{http://dx.doi.org/10.6028/NIST.TN.1838}}
\renewcommand{\headrulewidth}{0pt}

\begin{document}
\pagenumbering{gobble}

\bibliographystyle{unsrt}
%\pagestyle{empty}

@@

\begin{minipage}[t][9in][s]{6.25in}
\pagenumbering{gobble}

@@

\begingroup
This publication is available free of charge from:
\hypersetup{urlcolor=black}
\href{http://dx.doi.org/10.6028/NIST.TN.1838}{http://dx.doi.org/10.6028/NIST.TN.1838} \\
\endgroup

@@

\begin{minipage}[t][9in][s]{6.25in}
\pagenumbering{gobble}

@@

%\pagestyle{plain}
\pagenumbering{roman}
\setcounter{page}{3}

@@

\chapter*{\centering Abstract}
\pagenumbering{gobble}
The Fire Dynamics Simulator (FDS) fire model, which is developed and maintained by the National Institute of Standards and Technology (NIST), was used to provide insight into the dynamics of a fire that occurred on November 2, 2012, within a 2~\sfrac{1}{2}~story, single family structure in Chicago, IL, that resulted in the death of the captain of a fire department. The inputs for the FDS simulations are documented in this report and are based on the fire scenario, including the building geometry, interior furnishings, and ventilation conditions. The fire started in the attic and spread down through an opening into the enclosed porch on the rear of the structure. This resulted in ventilation limited (fuel rich) fire conditions in the attic and rear porch areas. The spread of fire and hot gases into the second floor of the structure was limited by a closed steel-faced, wood-framed door. After exposure to the elevated temperatures and pressure from the fire in the porch area, the wood frame of the door decomposed, the steel faces of the door failed, and the door collapsed inward. The door failure resulted in the establishment of a flow path between the higher pressure and higher temperature conditions in the enclosed porch and the lower pressure and lower temperature conditions in the hallway and kitchen areas of the second floor. The temperature of the gases that flowed into the hallway exceeded 260~$^{\circ}$C (500~$^{\circ}$F) at a height of 1.86~m (6~ft). Unfortunately, a captain and firefighter were advancing a hose line in the hallway near the door at the time of the door failure. It appears that the captain was likely overcome by the rapid change in conditions before he could exit to the safety of the kitchen area. After a Mayday call, the captain was rescued from the structure, but later succumbed to his injuries.

\chapter{Introduction}
\pagenumbering{arabic}
\setcounter{page}{1}
Part of the function of the Fire Research Division at NIST is to develop and apply technology, measurements, and standards to improve the understanding of the behavior, prevention, and control of fires to enhance fire fighting operations and equipment, fire suppression, fire investigations, and disaster response. NIST has previously used the Fire Dynamics Simulator (FDS) fire model to provide insight into the fire development and thermal conditions of fires that have resulted in line of duty deaths (LODDs)~\cite{Madrzykowski:1,Iowa,Texas,Bryner:Charleston,barowy:texas}. The objective of these studies has been to improve the safety and effectiveness of firefighters.

@@

\begin{table}
\centering
\captionof{table}{Relevant fire model input parameters}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter                                                  & Description                                   & Discussion                            \\
\midrule
Simulation Time                                            &  5~min                                        & --                                    \\ [.25cm]
Grid Cell Size                                             &  10~cm                                        & Section~\ref{mesh}                    \\ [.25cm]
Ambient Temperature$^*$                                    &  20~$^{\circ}$C (68~$^{\circ}$F)              & --                                    \\ [.1cm]
\multirow{4}{*}{Reaction: Wood~\cite{SFPE:Tewarson}}       &  Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$     & \multirow{4}{*}{Section~\ref{fire}}   \\
                                                           &  CO Yield: 0.004~$\mathrm{kg}/\mathrm{kg}$    &                                       \\
                                                           &  Soot Yield: 0.015~$\mathrm{kg}/\mathrm{kg}$  &                                       \\
                                                           &  Heat of Combustion: 16,400~kJ/kg             &                                       \\ [.25cm]
\multirow{2}{*}{Peak HRR}                                  &  Attic Fire: 9~MW                             & \multirow{2}{*}{Section~\ref{fire}}   \\ 
                                                           &  Porch Fire: 3.3~MW                           &                                       \\ [.25cm]                     
\multirow{3}{*}{Material: Wood~\cite{Incropera:1} }        &  $k$: 0.12~\si{W/(m.K)})                      & \multirow{3}{*}{Section~\ref{matl}}   \\
                                                           &  $\rho$: 510~kg/m$^3$                         &                                       \\
                                                           &  $c_{p}$: 1.38~\si{kJ/(kg.K)}                 &                                       \\ [.25cm]
\multirow{3}{*}{Material: Gypsum Board~\cite{WAKILI2007}}  &  $k$: 0.28~\si{W/(m.K)}                       &  \multirow{3}{*}{Section~\ref{matl}}  \\ 
                                                           &  $\rho$: 810~kg/m$^3$                         &                                       \\ 
                                                           &  $c_{p}$: 1.0~\si{kJ/(kg.K)}                  &                                       \\
\bottomrule[1.25pt]
\end{tabular}\par
\footnotesize
$^{*}$Initial interior temperatures were assumed to be 20~$^{\circ}$C; exterior temperatures were 7~$^{\circ}$C~\cite{NIOSH:Bowyer}.
\normalsize
\end{table}
